COLLISION RATES OF PLANETESIMALS NEAR MEAN-MOTION RESONANCES

### 2020-05-21 main.tex
Edit text, fix typos, add simulation names to figures

Recent observations of circumstellar disks by ALMA have revealed a rich variety of substructure in the millimeter contiunuum emission. Features 
such as gaps and asymmetries 
\citep{2015ApJ...808L...3A, 2016Sci...353.1519P, PhysRevLett.117.251101, 2016ApJ...820L..40A, 2016Natur.535..258C} in the emission provide 
diagnostics for the physical processes that drive the evolution of the disks. In some cases, these gap features are argued to be an indicator for the 
presence of a giant planet, either embedded in the disk \citep{2015MNRAS.453L..73D} or orbiting externally. A giant perturber can influence the 
structure of the continuum emission in a number of ways. A misaligned giant planet can produce nonaxisymmetric features such as warps 
\citep{2001A&A...370..447A}. Highly eccentric perturbers can produce even more complicated structures through secular perturbations 
\citep{2014MNRAS.443.2541P, 2015MNRAS.448.3679P}. Mean motion resonances (MMRs) have been shown to open gaps as well 
\citep{2015ApJ...798...83N, 2016ApJ...818..159T, 2018ApJ...857....3T}.

Collisions between planetesimals are thought to be a significant source of dust for many young circumstellar disks 
\citep[see][]{2008ARA&A..46..339W}. An expanding debris cloud from a planetesimal collision was recently detected through scattered light in the 
Fomalhut system \citep{2020PNAS..117.9712G}. Although some amount of primordial dust is likely still present during the planet formation process, 
ongoing collisions between small bodies will augment this and could be used to trace the underlying dynamical activity of the planetesimals. 
\citet{2013ApJ...777L..31D} showed that collisional dust generated near the gap formed by a giant planet in a planetesimal disk should produce a 
distinctly observable marker that could be used to infer the presence of the planet. In a followup study, it was found that morphological differences in 
the dust emission could be used to distinguish between a low and high eccentricity planet \citep{2016ApJ...820...29D}.

Although more difficult to spatially resolve, substructure due to mean-motion resonances (MMRs) with a giant planet may hold additional clues which 
could be used to constrain the properties of the planet.  The width of a resonance is set by both the mass of the perturbing planet, along with the 
unperturbed eccentricity of the planetesimals, which is in turn set by the eccentricity of the planet through secular forcing. Therefore, the structure of 
second generation dust produced near MMRs may hold information about both the mass and eccentricity of the planet.

@@

This work is organized in the following way: In section \ref{sec:dynamics}, we provide an overview of the relevant dynamics that drive the evolution 
of a planetesimal disk under the gravitational influence of an external perturber. In section \ref{sec:sims} we provide an overview of the N-body code 
that we use and describe the initial conditions that are used for five simulations in which a perturbing giant planet is given is given various masses 
and eccentricities. Section \ref{sec:results} presents the results of these five simulations and we use the collision statistics to generate synthetic dust 
emission profiles that would be detected with observing facilities like ALMA. Next, we show in section \ref{sec:constrain} that a characteristic bump 
or dip feature appears in the dust emission near the interior 2:1 MMR, the presence of which depends on the mass and eccentricity of the perturbing 
planet. We discuss the potential of using this feature to place and upper or lower on the mass and eccentricity of the planet and conclude in section 
\ref{sec:conclusions}.

@@

The most direct and widespread effect that a giant perturber will have on a planetesimal disk is through secular forcing of the 
eccentricities of the planetesimals. This will cause the complex eccentricities of the planetesimals to take on a time independent 
forced value, given by \citep{1999ApJ...527..918W} as

\begin{equation}\label{eq:eforced}
	z_{f} = \frac{b^{2}_{3/2} (\alpha)}{b^{1}_{3/2} (\alpha)} e_{g} ~ \mathrm{exp} ~ i \omega_{g}.
\end{equation}

\noindent Here, $\alpha = a_{g} / a$ where $a_{g}$ and $a$ are the semi-major axes of the giant planet and the planetesimal, 
respectively. $e_{g}$ and $\omega_{g}$ are the eccentricity and longitude of pericenter of the giant and $b^{j}_{s} (\alpha)$ is a 
Laplace coefficient given by \citep{2000ssd..book.....M} as

@@

If we further assume that the two bodies are orbiting in the same plane, the behavior of the bodies near resonance is determined by the
behavior of a critical angle

@@

\noindent where $\lambda = \varpi + M$ is the mean longitude of a body. For bodies in resonance, the critical angle will librate around an 
equlibrium value, while this angle will circulate outside of resonance. This behavior is analogous to the motion of a pendulum. Changes in the 
critical angle are coupled to changes in the mean motion and semimajor axis \citet{2000ssd..book.....M}. An important point to note, which will 
become important later, is that the circulation frequency approaches zero near the edge of a resonance. The width of a resonance can be 
defined by determining the largest variation in semimajor axis that permits librational, rather than circulatory motion of the pendulum. For 
second and higher order resonances, the maximum libration width is given by

@@

The net result of this is that MMRs will produce 'spikes' in the semimajor axis - eccentricity distribution of a planetesimal disk. In order to 
preserve the Jacobi energy, equation \ref{eq:tiss} predicts that these spikes will tend to curve away from the perturbing body. This lowers the 
surface density of the disk near the locations of resonances, while producing a pileup of bodies on eccentric orbits adjacent to the MMR. The 
enhanced surface density will increase the collision rate. At the same time, the large eccentricites will either lower it due to the decreased 
effectiveness of gravitational focusing, or raise it due to the more vigorous encounter rate, depending on how large the encounter velocity gets 
compared to the escape velocity of the planetesimals. In addition, the MMRs cause the orbits of planetesimals to precess on relatively short 
timescales, which effectively randomizes their longitudes of perihelion. In a secularly forced disk, there will be an interface between aligned 
and randomly oriented orbits at the edge of the resonance. Due to the complicated dynamics that are at play in this region, we turn to a direct 
N-body treatment to better understand how the collision rate varies in the vicinity of a MMR.

@@

In total, five simulations are run. The first is a 'nominal' case, in which the perturbing planet's mass and eccentricity are set to that of Jupiter's. 
In the other four cases, the mass (eccentricity) is increased or decreased by a factor of two, while the mass (eccentricity) is held at the nominal 
value. In all cases, the perturbing giant is placed on a 5.2 AU orbit around a 1 $M_{\odot}$ star. The planetesimal disk extends from 2.2 to 3.8 
AU, which covers the two most prominent mean motion resonances with the giant planet, the 2:1 at 3.27 AU and the 3:1 at 2.5 AU.The 
planetesimal disk follows a minimum-mass solar nebula surface density profile \citep{1981PThPS..70...35H}

@@

The dynamical effects of secular forcing by the giant planet, along with  the effects of viscous stirring and gas drag on the planetesimals, are 
built into the initial conditions. This is done by first calculating the equilibirum eccentricity $e_{eq}$ due to viscous stirring and gas drag as a 
function of semimajor axis according to equation 12 of \citet{2002ApJ...581..666K}. Although {\sc ChaNGa} does not account for the effects of 
gas drag on the planetesimals, the viscous stirring timescale is long compared to our integration time and should not cause the velocity 
distribution to significantly evolve. The eccentricities of the bodies are drawn randomly from a Rayleigh distribution with a mode of $e_{eq}$, 
while the inclinations are drawn from a similar distribution with a mode of $e_{eq}$/2 \citep{1993MNRAS.263..875I}. The longitude of perihelia 
$\omega$ and the mean anomalies $M$ of the bodies are drawn from a uniform polar distribution.

To account for the effects of secular forcing, the eccentricity vectors of the planetesimals are first decomposed into real and imaginary 
components:

\begin{equation}\label{eq:hk}
	z = (h, ik) = e \, exp(i \omega)
\end{equation}

\noindent and a forced component is added to $h$ according to equation \ref{eq:eforced} (where we have set $\omega_{g}$ = 0).

@@

For the purposes of the integrator, there are two relevant timescales in this system. The first is the orbital dynamical time $\sqrt{a^3/
G M_{\odot}}$. All particles are placed on a fixed time step of $\Delta T$ = 0.01 yr. This is roughly 3\% of an orbital dynamical time at 
the inner edge of the planetesimal disk. From numerical tests with {\sc ChaNGa}, we have found that this time step size preserves 
the symplectic nature of the integration without being too computationally expensive. Although using $\Delta T$ = 0.01 yr keeps the 
integration symplectic, we find that a small amount of artificial precession is produced. We reduce the base time step size by an additional 
factor of 4 to the keep longitude of perihelia of planetesimals near the inner edge of the disk from drifting too far over the course of our 
integrations.

@@

All five simulations are evolved for 5,000 years, which is roughly 400 orbits of the perturbing planet. Because the effects of the mean motion 
resonances are not built into the initial conditions, the simulations must be run long enough for the distribution of orbital elements to reach 
equilibrium near the resonances before the collision rate is measured. The libration timescale, which describes how long the critical angle (and also 
the semimajor axis and eccentricity) of a planetesimal in resonance takes to undergo a full oscillation is given by

@@

\noindent for small-amplitude librations. For all five sets of initial conditions used, $T_{lib} \sim$ 1,000 - 2,000 years for the 3:1 and 2:1 resonances. 
For this reason, we allow each simulation to run for 2,000 years before we begin tracking any collision statistics.

@@

We begin by examining simulations e1m2, e2m2 and e3m2. The positions of the planetesimals in the x-y plane after 5,000 years of integration is 
shown in figure \ref{fig:xy}. In all cases, the coordinate system is rotated so that the giant planet lies at $\theta = 0$ and the longitude of perihelion of 
the planet is shown by the dashed line. Resonances with the perturbing planet are visible via nonaxisymmetric gaps. Upon close inspection of the 
last few simulation snapshots, these gaps appear to follow the planet in its orbit, rather than aligning themselves with the longitude of perihelion. A 
similar substructure reveals itself in \citet{2000Icar..143...45R} (see bottom panel of figure 3) and \citet{2016ApJ...818..159T} (figure 3). It is worth 
noting that \citet{2000Icar..143...45R} started with a completely cold planetesimal disk and a Jupiter mass planet on a circular, coplanar orbit. The 
presence of these features seems robust to the choice of initial conditions.

@@

Next, we examine the statistics of collisions resolved in each of the simulations. The 3D positions and velocities of the two colliding bodies are 
recorded to a table at the moment of impact. From these quantities, we also derive the Keplerian orbital elements of a collision from these quantities. 
First, we examine the semimajor axis of the first particle participating in each collision. During a collision, the 'first' particle is defined as the more 
massive of the two. However, nearly all of the collisions happen between the initial, equal-mass planetesimals. In this case, the distinction is set by 
the collision search algorithm and is rather arbitrary. In all of the plots where we show collision statistics, we have verified that using the 'first' or 
'second' collider does not qualitatively change any of the features.

Figure \ref{fig:coll_hist_a} shows the semimajor axis distribution of the resolved collisions. We find that some of the features present in this and 
subsequent figures are highly sensitive to the number of bins and the location of the bin edges. For this reason, we construct a PDF of the collisions 
using a Kernel Density Estimate (KDE). We use the {\sc neighbors.KernelDensity} function from the {\sc sklearn} \citep{scikit-learn} package to 
construct our KDEs. For the kernel, we use a tophat function with a bandwidth of 0.02 AU.

Near some of the resonances, there are noticeable suppressions or enhancements of the local collision rate. This contrasts with the findings of 
\citet{2000Icar..143...45R}, who simulated a similar setup and found no discernible features near the MMRs. We attribute the differences to  a more 
conservative timestepping criterion in our simulations. The most prominent features appear as an enhancement to the collision rate near the 2:1 
MMR and a decrease near the 3:1 MMR. At higher forced eccentricities, features near the 5:2, 7:3 and 5:3 resonances are also visible, due to the 
steeper sensitivity of higher order resonances to eccentricity \citep{1994PhyD...77..289M}.

We attribute the production of a bump vs a dip at certain resonances to changes in the gravitational focusing factor during two-body encounters. 
Closer to the star, the planetesimal disk starts off dynamically colder. Here, additional heating from the resonance acts to suppress the gravitational 
focusing cross section, which reduces the collision rate from the non-perturbed value. Further out in the disk, where the encounter velocity is already 
higher, the gravitational focusing cross section is already minimized. Any additional heating instead acts to simply increase the encounter rate 
between bodies, which enhances the collision rate. This explains the relative drop in the collision rate near the 3:1 MMR and the bumps that form 
near the 2:1 and 5:3 resonance. Although this could potentially serve as a useful diagnostic of the planetesimal size in a disk (because the mutual 
escape velocity sets how dynamically hot the bodies can get before gravitational focusing is suppressed), a direct measurement of the semimajor 
axes of the planetesimals is not possible. In addition, as will demonstrate next, collisions due to bodies in resonance end up having an insignificant 
effect on the final shape of the radial dust distribution.

@@

\noindent where $m$, $e$ and $s$ are the mass, eccentricity and radius of the dust grain. $C_{D}$ is a drag coefficient which is of order unity, $
\rho_{g}$ is the local density of the gas and $v_{th}$ is its thermal speed. At 3 AU, the gaseous component of the solar nebula has a density of $4 
\times 10^{-10}$ g cm$^{-3}$ and a typical thermal speed of $10^{5}$ cm s$^{-1}$ \citep{1981PThPS..70...35H}. For a 1 mm dust grain with a 
density of 1 g cm$^{-3}$ and an eccentricity of 0.05, the damping timescale is $4 \times 10^{-3}$ years. This is orders of magnitude smaller than the 
collision timescale, therefore we conclude that collisionally generated dust grains will immediately couple to the gas.

Operating under this assumption, a map of the relative concentration of second-generation dust can be constructed from the cylindrical 
distance at which the collisions occur. This is shown in figure \ref{fig:coll_hist_r}. Similar to figure \ref{fig:coll_hist_a}, we use a KDE with a bandwidth 
of 0.02 AU to assemble the collision locations into a radial distribution. Most strikingly, the bump that was present near the 2:1 resonance is no 
longer present in cylindrical distance space, although it suddenly appears again in the e3m2 simulation. The dip that is present near the 3:1 MMR 
also disappears and instead presents as a bump in the e3m2 case. To determine how the resonant bodies actually contribute to the radial collision 
profile, we excluded collisions that fall between $2.495 < a < 2.505$ AU (near the 3:1 resonance) and between $3.2 < a < 3.35$ AU (near the 2:1 
resonance), which is shown by the orange curve. Qualitatively, none of the bump or dip features present are changed by making this exclusion. This 
suggests that the features near resonance seen in semimajor axis space become 'smeared out' in cylindrical distance space due to the large spread 
in eccentricity and orbital orientation.

This also suggests that the features seen in figure \ref{fig:coll_hist_r} are being produced by bodies outside of resonance. To further understand this, 
we revisit the nonaxisymmetric structure seen in figure \ref{fig:xy}. In figure \ref{fig:coll_polar_e}, we compare the radial collision profile to the polar 
structure seen in the e1m2, e2m2 and e3m2 simulations. In the lowest forced eccentricity case, the pileups near the edges of the resonances line up 
with the boundaries of the dip feature seen in the radial collision profile. The locations of the edges of the resonances are calculated using equations 
\ref{eq:res_fo} and \ref{eq:res_so}. The pileups are especially prominent for the 2:1 resonance. As discussed previously, the pileups and gaps seen 
in polar coordinates follow the position of the giant planet in its orbit.

Although the nonaxisymmetric pileup of bodies near the edges of resonances has been seen previously 
\citep{2000Icar..143...45R, 2016ApJ...818..159T}, we could not find a satisfactory explanation for why it happens and will provide one here. As 
mentioned previously, the circulation frequency of the critical angle slows as one approaches the edge of the resonance from the outside. Following 
the pendulum analogy, this is equivalent to reaching the point where the pendulum becomes suspended in the vertical position. When the critical 
angle stays relatively stationary during an encounter, the net torque will drive the longitude of pericenter towards the point of conjunction 
\citep{1976ARA&A..14..215P}.  This is the basic mechanism that causes isolated resonances to be stable. Inside of resonance, the critical angles 
librates about some value and this configuration is maintained. Just outside of resonance, however, the critical angle slowly drifts away, with the drift 
direction changing sign across the resonance. The net result of this is that bodies near the edge of the resonance experience a torque from that 
planet that temporarily drives their pericenters toward the current orbital position of the planet. After the encounter ends, differential rotation of the 
disk slowly destroys the configuration over part of an orbit. This explains why the concentration of planetesimals appears to 'follow' the the planet.

@@

\citet{2016ApJ...818..159T} provides a different explanation for this phenomenon. Instead, the nonaxisymmetric structure is claimed to be a product 
of the path that a resonant test particle takes in a frame co-rotating with the planet (see figure 8.4 of \citet{2000ssd..book.....M}). This explanation, 
however, does not hold for a large collection of planetesimals, especially when there is no forced eccentricity. In such a case, the orbits of 
planetesimals near resonance would be randomly aligned, and the axisymmetric structure should become washed out. In the case with a perturbing 
planet planet on a circular orbit, this structure is still present (see figure 3 of \citet{2016ApJ...818..159T}).

With no secular forcing, this dynamical phenomena would produce an underdensity in the radial surface density profile near the center of the 
resonance and an overdensity at each edge in cylindrical distance space. When a forced eccentricity is introduced, the edges of the resonance 
(indicated by the solid and dashed lines in figure \ref{fig:coll_polar_e}), where the pileups are located, begin to explore a wider range of radii over the 
course of a single orbit. If the the aphelion distance of the inner edge becomes comparable to the perihelion distance of the outer edge, the 
overdense regions on each side of the resonance meet and we should expect to see the central dip feature in the collision profile disappear. This is 
exactly what appears to be happening in the middle panel of figure \ref{fig:coll_polar_e}. As the forced eccentricity is increased further, a region 
forms at the center of the resonance where planetesimals from both sides of the resonance spend time (although not simultaneously). When this 
occurs, a bump feature forms in the collision profile near the center of the resonance. This is apparent in the bottom panel of figure 
\ref{fig:coll_polar_e}.

>>>>

Recent observations of circumstellar disks by ALMA have revealed a rich variety of substructure. Features 
such as gaps and asymmetries 
\citep{2015ApJ...808L...3A, 2016Sci...353.1519P, PhysRevLett.117.251101, 2016ApJ...820L..40A, 2016Natur.535..258C} in the emission provide 
diagnostics for the physical processes that drive the evolution of the disks. In some cases, these gap features are argued to be an indicator for the 
presence of a giant planet, either embedded in the disk \citep{2015MNRAS.453L..73D} or orbiting externally. A giant perturber can influence the 
structure of the continuum emission in a number of ways. A misaligned giant planet can produce nonaxisymmetric features such as warps 
\citep{2001A&A...370..447A}. Highly eccentric perturbers can produce even more complicated structures through secular interactions 
\citep{2014MNRAS.443.2541P, 2015MNRAS.448.3679P}. Mean motion resonances (MMRs) have been shown to open gaps as well 
\citep{2015ApJ...798...83N, 2016ApJ...818..159T, 2018ApJ...857....3T}.

Collisions between planetesimals are thought to be a significant source of dust for many young circumstellar disks 
\citep[see][]{2008ARA&A..46..339W}. An expanding debris cloud from a planetesimal collision was recently detected in the 
Fomalhut system \citep{2020PNAS..117.9712G}. Although some amount of primordial dust is likely still present during the planet formation process, 
ongoing collisions between small bodies will augment this and could be used to trace the underlying dynamical activity of the planetesimals. 
\citet{2013ApJ...777L..31D} showed that collisional dust generated near the gap formed by a giant planet in a planetesimal disk should produce a 
distinctly observable marker that could be used to infer the presence of the planet. In a followup study, it was found that morphological differences in 
the dust emission could be used to determine whether the planet had a circular or eccentric orbit \citep{2016ApJ...820...29D}.

Substructure due to mean-motion resonances (MMRs) with a giant planet may hold additional clues which 
could be used to constrain the properties of the planet.  The width of a resonance is set by both the mass of the perturbing planet, along with the 
unperturbed eccentricity of the planetesimals, which is in turn set by the eccentricity of the planet through secular forcing. Therefore, the structure of 
second generation dust produced near MMRs encodes information about both the mass and eccentricity of the planet.

@@

This work is organized in the following way: In section \ref{sec:dynamics}, we provide an overview of the relevant dynamics that drive the evolution 
of a planetesimal disk under the gravitational influence of an external perturber. In section \ref{sec:sims} we provide an overview of the N-body code 
used and describe the initial conditions chosen for five simulations in which a perturbing giant planet is given various masses 
and eccentricities. Section \ref{sec:results} presents the results of these simulations and we use the collision statistics to generate synthetic dust 
emission profiles that would be detected with observing facilities like ALMA. Next, we show in section \ref{sec:constrain} that a characteristic bump 
or dip feature appears in the dust emission near the interior 2:1 MMR, the presence of which depends on the mass and eccentricity of the perturbing 
planet. We discuss the potential of using this feature to place and upper or lower on the mass and eccentricity of the planet and conclude in section 
\ref{sec:conclusions}.

@@

The most direct and widespread effect that a giant perturber will have on a planetesimal disk is through secular forcing of the 
eccentricities of the planetesimals. This will cause the complex eccentricities of the planetesimals to take on a time-independent 
forced value, given by \citep{1999ApJ...527..918W} as

\begin{equation}\label{eq:eforced}
	z_{f} = \frac{b^{2}_{3/2} (\alpha)}{b^{1}_{3/2} (\alpha)} e_{g} ~ \mathrm{exp} ~ i \varpi_{g}.
\end{equation}

\noindent Here, $\alpha = a_{g} / a$ where $a_{g}$ and $a$ are the semi-major axes of the giant planet and the planetesimal, 
respectively. $e_{g}$ and $\varpi_{g}$ are the eccentricity and longitude of pericenter of the giant and $b^{j}_{s} (\alpha)$ is a 
Laplace coefficient given by \citep{2000ssd..book.....M} as

@@

If we further assume that the two bodies are orbiting in the same plane, the motion of the bodies near resonance is determined by the
behavior of a critical angle

@@

\noindent where $\lambda = \varpi + M$ is the mean longitude of a body. For bodies in resonance, the critical angle will librate around an 
equlibrium value, while this angle will circulate outside of resonance. For small eccentricities, this behavior is analogous to the motion of a 
pendulum. Variations in the critical angle are coupled to changes in the mean motion and semimajor axis \citet{2000ssd..book.....M}. An important 
point to note, which will become important later, is that the variation frequency of this angle approaches zero near the edge of a resonance. The 
width of a resonance can be defined by determining the largest variation in semimajor axis that permits librational, rather than circulatory motion of 
the $\phi$. For second and higher order resonances, the maximum libration width is given by

@@

The net result of this is that MMRs will produce spikes in the semimajor axis - eccentricity distribution of a planetesimal disk. In order to 
preserve the Jacobi energy, equation \ref{eq:tiss} predicts that these spikes will tend to curve away from the perturbing body. This lowers the 
surface density of the disk near the locations of resonances, while producing a pileup of bodies on eccentric orbits adjacent to the MMR. The 
enhanced surface density will increase the collision rate. At the same time, the large eccentricites will either lower it due to the decreased 
effectiveness of gravitational focusing, or raise it due to the more vigorous encounter rate, depending on the size of the encounter velocity relative to 
the escape velocity of the planetesimals. In addition, the MMRs cause the orbits of planetesimals to precess on relatively short 
timescales, which effectively randomizes their longitudes of perihelion. In a secularly forced disk, there will be an interface between aligned 
and randomly oriented orbits at the edge of the resonance. Due to the complicated dynamics that are at play in this region, we turn to a direct 
N-body treatment to better understand how the collision rate varies in the vicinity of a MMR.

@@

In total, five simulations are run. The first is a 'nominal' case, in which the perturbing planet's mass and eccentricity are set to that of Jupiter's. 
In the other four cases, the mass (eccentricity) is increased or decreased by a factor of two, while the eccentricity (mass) is held at the nominal 
value. In all cases, the perturbing giant is placed on a 5.2 AU orbit around a 1 $M_{\odot}$ star. The planetesimal disk extends from 2.2 to 3.8 
AU, which covers the two most prominent mean motion resonances with the giant planet, the 2:1 at 3.27 AU and the 3:1 at 2.5 AU. The 
planetesimal disk follows a minimum-mass solar nebula surface density profile \citep{1981PThPS..70...35H}

@@

The dynamical effects of secular forcing by the giant planet, along with  the effects of viscous stirring and gas drag on the planetesimals, are 
built into the initial conditions. This is done by first calculating the equilibirum eccentricity $e_{eq}$ due to viscous stirring and gas drag as a 
function of semimajor axis according to equation 12 of \citet{2002ApJ...581..666K}. Although {\sc ChaNGa} does not account for the effects of 
gas drag on the planetesimals, the viscous stirring timescale is long compared to our integration time and should not cause the velocity 
distribution to significantly evolve. The eccentricities of the bodies are drawn randomly from a Rayleigh distribution with a mode of $e_{eq}$, 
while the inclinations are drawn from a similar distribution with a mode of $e_{eq}$/2 \citep{1993MNRAS.263..875I}. The arguments of perihelion 
$\omega$, longitude of ascending nodes $\Omega$ and the mean anomalies $M$ of the bodies are drawn uniformly $\in [0, 2 \pi)$.

To account for the effects of secular forcing by the planet, the eccentricity vectors of the planetesimals are first decomposed into real and imaginary 
components:

\begin{equation}\label{eq:hk}
	z = (h, ik) = e \, exp(i \varpi)
\end{equation}

\noindent and a forced component is added to $h$ according to equation \ref{eq:eforced} (where we have set $\varpi_{g}$ = 0).

@@

For the purposes of the integrator, there are two relevant timescales in this system. The first is the orbital dynamical time $\sqrt{a^3/
G M_{\odot}}$. All particles are placed on a fixed time step of $\Delta T$ = 0.01 yr. This is roughly 3\% of an orbital dynamical time at 
the inner edge of the planetesimal disk. From numerical tests with {\sc ChaNGa}, we have found that this time step size preserves 
the symplectic nature of the integration without being too computationally expensive. Although using $\Delta T$ = 0.01 yr keeps the 
integration symplectic, we find that a small amount of artificial precession is produced. This is most noticeable at the inner edge of the
planetesimal disk. We reduce the base time step size by an additional 
factor of 4 to the keep longitude of perihelia of planetesimals near the inner edge of the disk from drifting too far over the course of our 
integrations.

@@

All five simulations are evolved for 5,000 years, which is roughly 400 complete orbits of the perturbing planet. Because the effects of the mean 
motion resonances are not built into the initial conditions, the simulations must be run long enough for the distribution of orbital elements to reach 
equilibrium near the resonances before the collision rate is measured. The libration timescale, which describes how long the critical angle (and also 
the semimajor axis and eccentricity) of a planetesimal in resonance takes to undergo a full oscillation is given by

@@

\noindent for small-amplitude librations. For all five sets of initial conditions used, $T_{lib} \sim$ 1,000 - 2,000 years for the 3:1 and 2:1 resonances. 
For this reason, we allow each simulation to run for 2,000 years before we begin recording any collision statistics.

@@

We begin by examining simulations e1m2, e2m2 and e3m2. The positions of the planetesimals in the x-y plane after 5,000 years of integration are 
shown in figure \ref{fig:xy}. In all cases, the coordinate system is rotated so that the giant planet lies at $\theta = 0$. The longitude of perihelion of 
the planet is shown by the dashed line. Resonances with the perturbing planet are visible as nonaxisymmetric gaps. Upon close inspection of a 
series of simulation snapshots, these gaps appear to follow the planet in its orbit, rather than aligning themselves with the longitude of perihelion. A 
similar substructure, which is most noticeable near the 2:1 MMR, reveals itself in \citet{2000Icar..143...45R} (see bottom panel of figure 3) and 
\citet{2016ApJ...818..159T} (figure 3). It is worth noting that \citet{2000Icar..143...45R} started with a completely cold planetesimal disk and a Jupiter 
mass planet on a circular, coplanar orbit. The presence of these features seems robust to the choice of initial conditions.

@@

Next, we examine the statistics of collisions resolved in each of the simulations. The 3D positions and velocities of the two colliding bodies are 
recorded to a table at the moment of impact. We derive the Keplerian orbital elements of a collision using these positions and velocities. 
First, we examine the semimajor axis of the first collider, which is shown in figure \ref{fig:coll_hist_a}. During a collision, the 'first' particle is defined 
as the more massive of the two. However, nearly all of the collisions happen between the initial, equal-mass planetesimals. In this case, the 
distinction is set by the collision search algorithm and is rather arbitrary. In all of the plots where we show collision statistics, we have verified that 
using the 'first' or 'second' collider does not qualitatively change any of the features. We find that some of the features present in figure 
\ref{fig:coll_hist_a} in this and subsequent figures are highly sensitive to the number of bins and the location of the bin edges. For this reason, we 
construct a PDF of the collisions using a Kernel Density Estimate (KDE). We use the {\sc neighbors.KernelDensity} function from the {\sc sklearn} 
\citep{scikit-learn} package to construct our KDEs. For the kernel, we use a tophat function with a bandwidth of 0.02 AU.

Near the stronger resonances, there are noticeable suppressions or enhancements to the local collision rate. This contrasts with the findings of 
\citet{2000Icar..143...45R}, who simulated a similar setup and found no discernible features near the MMRs. We attribute the differences to a more 
conservative timestepping criterion in our simulations. The most prominent features appear as an enhancement to the collision rate near the 2:1 
MMR and a decrease near the 3:1 MMR. At higher forced eccentricities, features near the 5:2, 7:3 and 5:3 resonances are also visible, due to the 
steep sensitivity of higher order resonant perturbations to eccentricity \citep{1994PhyD...77..289M}.

We attribute the production of a bump or a dip at certain resonances to changes in the gravitational focusing factor during two-body encounters. 
Closer to the star, the planetesimal disk starts off dynamically colder. Here, additional heating from the resonance acts to suppress the gravitational 
focusing cross section, which reduces the collision rate from its non-perturbed value. Further out in the disk, where the unperturbed encounter velocity is higher, the gravitational focusing cross section is already minimized. Any additional heating instead acts to simply increase the encounter rate between bodies, which enhances the collision rate. This explains the relative drop in the collision rate near the 3:1 MMR and the relative increase near the 2:1 and 5:3 resonances. Although this could potentially serve as a useful diagnostic of the planetesimal size in a disk (because the mutual escape velocity sets how dynamically hot the bodies can get before gravitational focusing is suppressed), a direct measurement of the semimajor axes of the colliding planetesimals is not possible. Furthermore, as will demonstrate next, collisions due to bodies in resonance do not have a significant effect on the final shape of the radial dust distribution.

@@

\noindent where $m$, $e$ and $s$ are the mass, eccentricity and radius of the dust grain. $C_{D}$ is a drag coefficient which is of order unity, $
\rho_{g}$ is the local density of the gas and $v_{th}$ is its thermal speed. At 3 AU, the gaseous component of the solar nebula has a density of $4 
\times 10^{-10}$ g cm$^{-3}$ and a typical thermal speed of $10^{5}$ cm s$^{-1}$ \citep{1981PThPS..70...35H}. For a 1 mm dust grain with a 
density of 1 g cm$^{-3}$ and an eccentricity of 0.05, the damping timescale is $4 \times 10^{-3}$ years. This is orders of magnitude smaller than the 
orbital timescale, therefore we conclude that collisionally generated dust grains will immediately couple to the gas.

Operating under this assumption, a map of the relative concentration of second-generation dust can be constructed from the cylindrical 
distance at which the collisions occur. This is shown in figure \ref{fig:coll_hist_r}. Similar to figure \ref{fig:coll_hist_a}, we use a KDE with a bandwidth 
of 0.02 AU to assemble the collision locations into a radial distribution. Most strikingly, the bump that was present near the 2:1 resonance is no 
longer visible in cylindrical distance space, although it suddenly appears again in the e3m2 simulation. The dip that is visible near the 3:1 MMR 
also becomes a bump in the highest eccentricity case. To determine how the resonant bodies actually contribute to the radial collision 
profile, we excluded collisions that fall between $2.495 < a < 2.505$ AU (near the 3:1 resonance) and between $3.2 < a < 3.35$ AU (near the 2:1 
resonance), which is shown by the orange curve. Qualitatively, none of the bump or dip features present are removed by making this exclusion. This 
suggests that the features near resonance seen in semimajor axis space become smeared out in cylindrical distance space due to the large spread 
in eccentricity and orbital orientation.

This also suggests that the most prominent bumps and dips seen in figure \ref{fig:coll_hist_r} are being produced by bodies outside of resonance. To 
further understand this, we consider the nonaxisymmetric structure seen in figure \ref{fig:xy}. In figure \ref{fig:coll_polar_e}, we compare the radial 
collision profile to the radial structure seen in the e1m2, e2m2 and e3m2 simulations. In the lowest forced eccentricity case, the pileups near the 
edges of the resonances line up with the boundaries of the dip feature seen in the radial collision profile. The locations of the edges of the 
resonances are calculated using equations \ref{eq:res_fo} and \ref{eq:res_so}. The pileups are especially prominent for the 2:1 resonance. As 
discussed previously, the pileups and gaps seen in polar coordinates follow the position of the giant planet in its orbit, which is shown by the vertical 
line.

Although the nonaxisymmetric pileup of bodies near the edges of resonances has been seen previously 
\citep{2000Icar..143...45R, 2016ApJ...818..159T}, we could not find a satisfactory explanation for why it happens and will provide one here. As 
mentioned in section \ref{sec:dynamics}, the circulation frequency of the critical angle slows as one approaches the edge of the resonance from the outside. Following the pendulum analogy, this is equivalent to approaching the point where the pendulum becomes suspended in the vertical position. When the critical angle stays relatively stationary during an encounter, the net torque will drive the longitude of pericenter towards the point of conjunction \citep{1976ARA&A..14..215P}.  This is the basic mechanism that causes isolated resonances to be stable. Inside of resonance, the critical angles librates about some value and this configuration is maintained. Just outside of resonance, however, the critical angle slowly drifts away, with the drift direction changing sign across the resonance. The net result of this is that bodies just outside of the resonance experience a torque from that planet that temporarily drives their pericenters toward the current orbital position of the planet. After the encounter ends, differential rotation of the disk slowly destroys the configuration over part of an orbit. As the planet passes by, adjacent planetesimals near the resonance edges temporarily align themselves in this configuration. This explains why the concentration of planetesimals appears to 'follow' the the planet.

@@

\citet{2016ApJ...818..159T} provides a different explanation for this phenomenon. Instead, the nonaxisymmetric structure is claimed to be a product 
of the path that a resonant test particle takes in a frame co-rotating with the planet (see figure 8.4 of \citet{2000ssd..book.....M}). This explanation, 
however, does not hold for a large collection of planetesimals, especially when there is no forced eccentricity. In such a case, the orbits of 
planetesimals near resonance would be randomly aligned, and the axisymmetric structure should become washed out. In the case with a perturbing 
planet planet on a circular orbit, this structure is still present (see figure 3 of \citet{2016ApJ...818..159T}). The only relation between the structure of 
the gaps in the disk and the path of a resonant particle in the corotating frame is that both phenomena produce the same symmetries, which depend 
on the resonance being considered.

With no secular forcing, this dynamical phenomena should produce an underdensity in the radial surface density profile near the center of the 
resonance and an overdensity at each edge in cylindrical distance space. When a forced eccentricity is introduced, the edges of the resonance 
(indicated by the solid and dashed lines in figure \ref{fig:coll_polar_e}), where the pileups are located, begin to explore a wider range of radii over the 
course of a single orbit. If the the aphelion distance of the inner edge becomes comparable to the perihelion distance of the outer edge, the 
overdense regions on each side of the resonance meet and we should expect to see the central dip feature in the collision profile disappear. This is 
exactly what appears to be happening in the middle panel of figure \ref{fig:coll_polar_e}. As the forced eccentricity is increased further, a region 
forms at the center of the resonance where planetesimals on both sides of the resonance spend time (although not simultaneously). When this 
occurs, a bump feature forms in the collision profile near the center of the resonance. This is apparent in the bottom panel of figure 
\ref{fig:coll_polar_e}.
